---
title: "RPS Bot Analysis"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(patchwork)
library(knitr)
```


```{r globals, include=FALSE}
DATA_FILE = "rps_v2_data.csv" # name of file containing full dataset for all rounds
FREE_RESP_FILE = "rps_v2_data_freeResp.csv" # file containing free response data by participant
SLIDER_FILE = "rps_v2_data_sliderData.csv" # file containing slider Likert data by participant
NUM_ROUNDS = 300 # number of rounds in each complete game
BLOCKSIZE = NUM_ROUNDS / 10 # number of bins to use for segmented analysis of performance over the full game
STRATEGY_LEVELS = c("prev_move_positive", "prev_move_negative",
                    "opponent_prev_move_positive", "opponent_prev_move_nil",
                    "win_nil_lose_positive", "win_positive_lose_negative",
                    "outcome_transition_dual_dependency")
STRATEGY_LOOKUP = list("prev_move_positive" = "Previous move (+)",
                       "prev_move_negative" = "Previous move (-)",
                       "opponent_prev_move_positive" = "Opponent previous move (+)",
                       "opponent_prev_move_nil" = "Opponent previous move (0)",
                       "win_nil_lose_positive" = "Win-stay-lose-positive",
                       "win_positive_lose_negative" = "Win-positive-lose-negative",
                       "outcome_transition_dual_dependency" = "Outcome-transition dual dependency")
```


```{r analysis_functions, echo=FALSE}
# Read in and process game results data (includes removal of duplicates, etc.)
read_game_data = function(filename) {
  data = read_csv(filename)
  data$bot_strategy = factor(data$bot_strategy, levels = STRATEGY_LEVELS)
  
  # Remove all incomplete games
  incomplete_games = data %>%
    group_by(game_id, player_id) %>%
    summarize(rounds = max(round_index)) %>%
    filter(rounds < NUM_ROUNDS) %>%
    select(game_id) %>%
    unique()

  data = data %>%
    filter(!(game_id %in% incomplete_games$game_id))
  
  # Remove any duplicate complete games that have the same SONA survey code
  # NB: this can happen if somebody played all the way through but exited before receiving credit
  # First, fetch sona survey codes with multiple complete games
  repeat_codes = data %>%
    group_by(sona_survey_code) %>%
    filter(is_bot == 0) %>%
    summarize(trials = n()) %>%
    filter(trials > 300) %>%
    select(sona_survey_code)
  
  # Next, get game id for the earlier complete game
  # NB: commented out code checks that we have slider/free resp data for at least one of the games
  duplicate_games = data %>%
    filter(sona_survey_code %in% repeat_codes$sona_survey_code &
             is_bot == 0  &
             round_index == NUM_ROUNDS) %>%
    select(sona_survey_code, game_id, player_id, round_begin_ts) %>%
    # remove the earlier one since the later one has free response and slider data (confirm with joins below)
    group_by(sona_survey_code) %>%
    filter(round_begin_ts == min(round_begin_ts)) %>%
    # inner_join(fr_data, by = c("game_id", "player_id")) %>%
    # inner_join(slider_data, by = c("game_id", "player_id")) %>%
    distinct(game_id)
  
  data = data %>%
    filter(!game_id %in% duplicate_games$game_id)
  
  return(data)
}

# Read in and process free response data
read_free_resp_data = function(filename, game_data) {
  data = read_csv(filename)
  # Join with game data to get bot strategy, etc.
  data = data %>%
    inner_join(game_data, by = c("game_id", "player_id")) %>%
    distinct(bot_strategy, game_id, player_id, free_resp_answer)
  # Order bot strategies
  data$bot_strategy = factor(data$bot_strategy, levels = STRATEGY_LEVELS)
  # Add plain english strategy, string process free resposne answers
  data = data %>%
    group_by(bot_strategy, player_id) %>%
    mutate(strategy = STRATEGY_LOOKUP[[bot_strategy]],
           free_resp_answer = str_replace_all(free_resp_answer, "\n" , "[newline]")) %>%
    ungroup()

  return(data)
}

# Read in and process slider data
read_slider_data = function(filename, game_data) {
  data = read_csv(filename)
  data = data %>%
    inner_join(game_data, by = c("game_id", "player_id")) %>%
    distinct(game_id, player_id, bot_strategy, index, statement, resp)
  # Order bot strategies
  data$bot_strategy = factor(data$bot_strategy, levels = STRATEGY_LEVELS)
  # Add plain english strategy
  data = data %>%
    group_by(bot_strategy, player_id, index) %>%
    mutate(strategy = STRATEGY_LOOKUP[[bot_strategy]]) %>%
    ungroup()
  
  return(data)
}

get_slider_summary = function(slider_data) {
  slider_data %>%
    group_by(statement, bot_strategy, strategy) %>%
    summarize(n = n(),
              mean_resp = mean(resp),
              se = sd(resp) / sqrt(n),
              se_upper = mean_resp + se,
              se_lower = mean_resp - se)
}

# Get win count differential by bot strategy, player
get_empirical_win_count_differential = function(data) {
  # NB: this is different from the empirical win count differential in v1 because
  # we care about human wins - bot wins, not absolute value between each player
  win_diff = data %>%
    group_by(bot_strategy, game_id, player_id, is_bot) %>%
    count(win_count = player_outcome == "win") %>%
    filter(win_count == TRUE) %>%
    group_by(bot_strategy, game_id) %>%
    summarize(win_count_diff = n[is_bot == 0] - n[is_bot == 1]) %>%
    as.data.frame()
  return(win_diff)
}

# Get win count differential for random behavior
get_null_win_count_differential = function(reps) {
  # NB: this differs slightly from the empirical null from v1 because
  # we care about humans *beating* bots, not absolute value between both opponents
  win_diff_null = data.frame(
    win_count_diff = replicate(reps, sum(sample(c(-1, 0, 1), NUM_ROUNDS, replace = T)))
  )
  return(win_diff_null)
}

# Get summary statistics for win count differential by strategy
get_win_count_differential_summary = function(strategy_data) {
  strategy_data %>%
    group_by(bot_strategy) %>%
    summarize(mean_win_count_diff = mean(win_count_diff),
              n = n(),
              se = sd(win_count_diff) / sqrt(n),
              lower_se = mean_win_count_diff - se,
              upper_se = mean_win_count_diff + se)
}

# Divide each subject's trials into blocks of size blocksize (e.g. 10 trials)
# then get each subject's win percent in each block
get_subject_block_data = function(data, blocksize) {
  data %>%
    filter(is_bot == 0) %>%
    group_by(bot_strategy, round_index) %>%
    mutate(round_block = ceiling(round_index / blocksize)) %>%
    select(bot_strategy, round_index, game_id, player_id, player_outcome, round_block) %>%
    group_by(bot_strategy, game_id, player_id, round_block) %>%
    count(win = player_outcome == "win") %>%
    mutate(total = sum(n),
           win_pct = n / total) %>%
    filter(win == TRUE)
}

# Take in subject block win percent (calculated above) and summarize by bot strategy across subjects
get_block_data_summary = function(subject_block_data) {
  subject_block_data %>%
    group_by(bot_strategy, round_block) %>%
    summarize(subjects = n(),
              mean_win_pct = mean(win_pct),
              se_win_pct = sd(win_pct) / sqrt(subjects),
              lower_ci = mean_win_pct - se_win_pct,
              upper_ci = mean_win_pct + se_win_pct)
}

# Get loss percent for each bot dependent on their previous move
get_bot_prev_move_loss_pct = function(data) {
  data %>%
    filter(bot_strategy == "prev_move_positive" | bot_strategy == "prev_move_negative") %>%
    group_by(player_id) %>%
    mutate(prev_move = lag(player_move, 1)) %>%
    filter(is_bot == 1, # look only at bot prev moves
           !is.na(prev_move), # lag call above sets NA for lag on first move: ignore it here
           prev_move != "none") %>%
    group_by(bot_strategy, game_id, player_id, prev_move) %>%
    count(player_outcome) %>%
    filter(!is.na(player_outcome)) %>% # TODO why do we have NA game outcomes??
    group_by(bot_strategy, game_id, player_id, prev_move) %>%
    # player win percent calculated as bot loss percent
    summarize(player_win_pct = n[player_outcome == "loss"] / sum(n))
}

# Get win percent for each player dependent on their own previous move
get_player_prev_move_win_pct = function(data) {
  data %>%
    filter(is_bot == 0,
           bot_strategy == "opponent_prev_move_nil" | bot_strategy == "opponent_prev_move_positive") %>%
    group_by(player_id) %>%
    mutate(prev_move = lag(player_move, 1)) %>%
    filter(!is.na(prev_move), # lag call above sets NA for lag on first move: ignore it here
           prev_move != "none") %>%
    group_by(bot_strategy, game_id, player_id, prev_move) %>%
    count(player_outcome) %>%
    filter(!is.na(player_outcome)) %>% # TODO why do we have NA game outcomes??
    group_by(bot_strategy, game_id, player_id, prev_move) %>%
    # player win percent calculated using player win outcomes
    summarize(player_win_pct = n[player_outcome == "win"] / sum(n))
}

# Get loss percent for each bot dependent on the bot's previous outcome
get_bot_prev_outcome_loss_pct = function(data) {
  data %>%
    filter(bot_strategy == "win_nil_lose_positive" | bot_strategy == "win_positive_lose_negative") %>%
    group_by(player_id) %>%
    mutate(prev_outcome = lag(player_outcome, 1)) %>%
    filter(is_bot == 1, # look only at bot prev moves
           !is.na(prev_outcome)) %>% # lag call above sets NA for lag on first move: ignore it here
    group_by(bot_strategy, game_id, player_id, prev_outcome) %>%
    count(player_outcome) %>%
    filter(!is.na(player_outcome)) %>% # TODO why do we have NA game outcomes??
    group_by(bot_strategy, game_id, player_id, prev_outcome) %>%
    # player win percent calculated as bot loss percent
    summarize(player_win_pct = n[player_outcome == "loss"] / sum(n))
}

# Get loss percent for each bot dependent on the bot's outcome two rounds ago
get_bot_prev2_outcome_loss_pct = function(data) {
  data %>%
    filter(bot_strategy == "win_nil_lose_positive" | bot_strategy == "win_positive_lose_negative") %>%
    group_by(player_id) %>%
    mutate(prev_outcome = lag(player_outcome, 2)) %>%
    filter(is_bot == 1, # look only at bot prev moves
           !is.na(prev_outcome)) %>% 
    group_by(bot_strategy, game_id, player_id, prev_outcome) %>%
    count(player_outcome) %>%
    filter(!is.na(player_outcome)) %>% # TODO why do we have NA game outcomes??
    group_by(bot_strategy, game_id, player_id, prev_outcome) %>%
    # player win percent calculated as bot loss percent
    summarize(player_win_pct = n[player_outcome == "loss"] / sum(n))
}

# Get summary win percent by strategy (dependent on previous move by bot or player)
get_prev_move_win_pct_summary = function(prev_move_data) {
  prev_move_data %>%
    group_by(bot_strategy, prev_move) %>%
    summarize(mean_player_win_pct = mean(player_win_pct),
              n = n(),
              se = sd(player_win_pct) / sqrt(n),
              se_lower = mean_player_win_pct - se,
              se_upper = mean_player_win_pct + se)
}


# Get summary win percent by strategy (dependent on previous outcome by bot or player)
get_prev_outcome_win_pct_summary = function(bot_loss_prev_outcome) {
  bot_loss_prev_outcome %>%
    group_by(bot_strategy, prev_outcome) %>%
    summarize(mean_player_win_pct = mean(player_win_pct),
              n = n(),
              se = sd(player_win_pct) / sqrt(n),
              se_lower = mean_player_win_pct - se,
              se_upper = mean_player_win_pct + se)
}

```



```{r graph_style, echo=FALSE}
default_plot_theme = theme(
  # titles
  plot.title = element_text(face = "bold", size = 18),
  axis.title.y = element_text(face = "bold", size = 16),
  axis.title.x = element_text(face = "bold", size = 16),
  legend.title = element_text(face = "bold", size = 16),
  # axis text
  axis.text.y = element_text(size = 14, face = "bold"),
  axis.text.x = element_text(size = 14, angle = 45, vjust = 0.5, face = "bold"),
  # legend text
  legend.text = element_text(size = 16, face = "bold"),
  # facet text
  strip.text = element_text(size = 12),
  # backgrounds, lines
  panel.background = element_blank(),
  strip.background = element_blank(),
  
  panel.grid = element_line(color = "gray"),
  axis.line = element_line(color = "black"),
  # positioning
  legend.position = "bottom",
  legend.key = element_rect(colour = "transparent", fill = "transparent")
)
```



```{r graphing_functions, echo=FALSE}
# Histogram of win count differentials for each dyad in a given strategy group
plot_win_count_differential = function(win_diff, group, min, max, labelx, labely) {
  avg = round(mean(win_diff$win_count_diff), 1)
  sd = round(sd(win_diff$win_count_diff), 1)
  win_diff %>%
    ggplot(aes(x = win_count_diff, color = group, fill = group)) +
    geom_histogram(alpha = 0.5, breaks = c(seq(min, max, by = 25))) +
    geom_text(aes(x = labelx, y = labely, label = str_c("Mean: ", avg, " (sd: ", sd, ")")), color = "red", size = 4) +
    labs(x = "win count differential", y = "count") +
    scale_color_viridis(discrete = T,
                        name = element_blank(),
                        begin = 0.2,
                        end = 0.8) +
    scale_fill_viridis(discrete = T,
                       name = element_blank(),
                       begin = 0.2, 
                       end = 0.8) +
    ggtitle(group) +
    default_plot_theme +
    theme(legend.position = "none")
}

# Histogram of null win count differential 
# NB: similar to above but with some formatting changes that make it easier to use a separate function
plot_null_win_count_differential = function(win_diff, group, min, max) {
  win_diff %>%
    ggplot(aes(x = win_count_diff, color = group, fill = group)) +
    geom_histogram(alpha = 0.5, breaks = c(seq(min, max, by = 5))) +
    labs(x = "win count differential", y = "count") +
    scale_color_viridis(discrete = T,
                        name = element_blank(),
                        begin = 0.2,
                        end = 0.8) +
    scale_fill_viridis(discrete = T,
                       name = element_blank(),
                       begin = 0.2, 
                       end = 0.8) +
    ggtitle(group) +
    default_plot_theme +
    theme(legend.position = "none")
}

# Plot mean + SEM of each strategy
plot_win_count_differential_summary = function(wcd_summary) {
  label_width = 10
  summary_labels = c("prev_move_positive" = str_wrap(STRATEGY_LOOKUP[["prev_move_positive"]], label_width),
                   "prev_move_negative" = str_wrap(STRATEGY_LOOKUP[["prev_move_negative"]], label_width),
                   "opponent_prev_move_positive" = str_wrap(STRATEGY_LOOKUP[["opponent_prev_move_positive"]], label_width),
                   "opponent_prev_move_nil" = str_wrap(STRATEGY_LOOKUP[["opponent_prev_move_nil"]], label_width),
                   "win_nil_lose_positive" = str_wrap(STRATEGY_LOOKUP[["win_nil_lose_positive"]], label_width),
                   "win_positive_lose_negative" = str_wrap(STRATEGY_LOOKUP[["win_positive_lose_negative"]], label_width),
                   "outcome_transition_dual_dependency" = str_wrap(STRATEGY_LOOKUP[["outcome_transition_dual_dependency"]], label_width))
  
  wcd_summary %>%
    ggplot(aes(x = bot_strategy, y = mean_win_count_diff)) +
    geom_point(aes(color = bot_strategy),
               size = 6) +
    geom_errorbar(aes(color = bot_strategy, ymin = lower_se, ymax = upper_se),
                  width = 0.25, size = 1) +
    geom_hline(yintercept = 0, size = 1, linetype = "dashed", color = "red") +
    labs(x = "", y = "mean win count differential") +
    ggtitle("Win count differential across bot strategies") +
    scale_x_discrete(labels = summary_labels) +
    scale_color_viridis(discrete = TRUE,
                        name = element_blank()) +
    default_plot_theme +
    theme(
      # plot.title = element_text(size = 32, face = "bold"),
      # axis.title.y = element_text(size = 24, face = "bold"),
      # axis.text.x = element_text(size = 20, face = "bold", angle = 0, vjust = 1),
      axis.text.x = element_text(angle = 0, vjust = 1),
      # axis.text.y = element_text(face = "bold", size = 20),
      legend.position = "none"
    )
}

# Plot average of each participant's win percent in blocks of trials by strategy
plot_win_pct_by_block = function(block_data_summary) {
  label_width = 20
  strategy_labels = c("prev_move_positive" = str_wrap(STRATEGY_LOOKUP[["prev_move_positive"]], label_width), 
                    "prev_move_negative" = str_wrap(STRATEGY_LOOKUP[["prev_move_negative"]], label_width),
                    "opponent_prev_move_nil" = str_wrap(STRATEGY_LOOKUP[["opponent_prev_move_nil"]], label_width),
                    "opponent_prev_move_positive" = str_wrap(STRATEGY_LOOKUP[["opponent_prev_move_positive"]], label_width),
                    "win_nil_lose_positive" = str_wrap(STRATEGY_LOOKUP[["win_nil_lose_positive"]], label_width),
                    "win_positive_lose_negative" = str_wrap(STRATEGY_LOOKUP[["win_positive_lose_negative"]], label_width),
                    "outcome_transition_dual_dependency" = str_wrap(STRATEGY_LOOKUP[["outcome_transition_dual_dependency"]], label_width))
  
  block_data_summary %>%
    ggplot(aes(x = round_block, y = mean_win_pct, color = bot_strategy)) +
    geom_point(size = 6, alpha = 0.75) +
    geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), size = 1, width = 0.25, alpha = 0.75) +
    geom_hline(yintercept = 1 / 3, linetype = "dashed", color = "red", size = 1) +
    labs(x = "Block of N rounds", y = "Mean win percentage") +
    ggtitle("Participant win percentage against bot strategies") +
    scale_color_viridis(discrete = T,
                        name = element_blank(),
                        labels = strategy_labels) +
    default_plot_theme +
    theme(axis.text.x = element_blank(),
          legend.text = element_text(face = "plain"))
}

# Plot average win percent based on previous move dependency
plot_prev_move_win_pct = function(bot_loss_summary_prev_move, strategy, xlabel) {
  bot_loss_summary_prev_move %>%
    filter(bot_strategy == strategy) %>%
    ggplot(aes(x = prev_move, y = mean_player_win_pct)) +
    geom_bar(stat = "identity", alpha = 0.5, color = "grey50", fill = "steelblue") +
    geom_errorbar(aes(ymin = se_lower, ymax = se_upper), width = 0.5, size = 1, color = "midnightblue") +
    geom_hline(yintercept = 1 / 3, linetype = "dashed", color = "red", size = 1) +
    labs(x = xlabel, y = "Avg. player win pct.") +
    ggtitle(STRATEGY_LOOKUP[[strategy]]) +
    default_plot_theme +
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 0, vjust = 1))
}

# Plot average win percent based on outcome dependency
plot_outcome_win_pct = function(bot_loss_summary_prev_outcome, strategy, xlabel) {
  bot_loss_summary_prev_outcome %>%
    filter(bot_strategy == strategy) %>%
    ggplot(aes(x = prev_outcome, y = mean_player_win_pct)) +
    geom_bar(stat = "identity", alpha = 0.5, color = "grey50", fill = "steelblue") +
    geom_errorbar(aes(ymin = se_lower, ymax = se_upper), width = 0.5, size = 1, color = "midnightblue") +
    geom_hline(yintercept = 1 / 3, linetype = "dashed", color = "red", size = 1) +
    labs(x = xlabel, y = "Avg. player win pct") +
    ggtitle(STRATEGY_LOOKUP[[strategy]]) +
    default_plot_theme +
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 0, vjust = 1))
}

# Plot mean + SEM of slider responses for a given question
plot_slider_data = function(slider_data) {
  label_width = 10
  strategy_labels = c("prev_move_positive" = str_wrap(STRATEGY_LOOKUP[["prev_move_positive"]], label_width), 
                      "prev_move_negative" = str_wrap(STRATEGY_LOOKUP[["prev_move_negative"]], label_width),
                      "opponent_prev_move_nil" = str_wrap(STRATEGY_LOOKUP[["opponent_prev_move_nil"]], label_width),
                      "opponent_prev_move_positive" = str_wrap(STRATEGY_LOOKUP[["opponent_prev_move_positive"]], label_width),
                      "win_nil_lose_positive" = str_wrap(STRATEGY_LOOKUP[["win_nil_lose_positive"]], label_width),
                      "win_positive_lose_negative" = str_wrap(STRATEGY_LOOKUP[["win_positive_lose_negative"]], label_width),
                      "outcome_transition_dual_dependency" = str_wrap(STRATEGY_LOOKUP[["outcome_transition_dual_dependency"]], label_width))
  q = unique(slider_data$statement)
  slider_data %>%
    ggplot(aes(x = bot_strategy, y = mean_resp, color = bot_strategy)) +
    geom_point(size = 6) +
    geom_errorbar(aes(ymin = se_lower, ymax = se_upper), size = 1, width = 0.25) +
    scale_color_viridis(discrete = T,
                        name = element_blank(),
                        labels = element_blank()) +
    scale_x_discrete(name = element_blank(),
                   labels = strategy_labels) +
    ylim(c(1, 7)) +
    labs(y = "Mean response (1: Strongly disagree, 7: Strongly agree)") +
    ggtitle(str_wrap(q, 50)) +
    default_plot_theme +
    theme(axis.text.x = element_text(angle = 0, vjust = 1),
          axis.title.x = element_blank(),
          legend.position = "none")
}
```



```{r analysis, message=FALSE, echo=FALSE} 
data = read_game_data(DATA_FILE)

participants = data %>%
  filter(is_bot == 0) %>%
  select(player_id) %>%
  unique() %>%
  summarize(n = n())

bots = length(unique(data$bot_strategy))
```



## Introduction

We investigate people's performance in the rock, paper, scissors game (RPS) against a series of bots employing increasingly complex strategies. 

What patterns do people detect when playing against bots that have a stable dependency structure? Is people's ability to win consistently against a bot predicted by the complexity of the bot's strategy?

We find that people show a reliable ability to exploit simple patterns based on their own or the bot's previous moves, but fail to comprehensively exploit more complex dependency structures that are conditioned on both previous moves and their accompanying outcomes, despite the fact that people's own move choices exhibit these dependencies when playing against other humans (Experiment 1).



## Methods 

A total of `r participants$n` participants played matches consisting of `r NUM_ROUNDS` rounds of RPS. Each participant was paired up against one of `r bots` stable bot strategies for the duration of the match.

The number of participants that played complete matches against each bot strategy is shown below. 


```{r participants, echo=FALSE}
data %>%
  filter(is_bot == 0) %>%
  group_by(bot_strategy) %>%
  summarize(n = n() / NUM_ROUNDS) %>%
  group_by(bot_strategy) %>%
  mutate(bot_strategy_plain = STRATEGY_LOOKUP[[bot_strategy]]) %>%
  ungroup() %>%
  select(bot_strategy_plain, n) %>%
  kable(col.names = c("Strategy", "N"))

```


In detail, the strategies in the table above are as follows:

* __Previous move (+)__: The bot has a 90% chance of making a *positive transition (+)* from its own previous move (i.e., rock -> paper, paper -> scissors, scissors -> rock). In other words, the bot usually transitions to the move that will beat what it just played (here and in all other strategies, the remaining 10% is distributed evenly across the other two transitions).

* __Previous move (-)__: The bot has a 90% chance of making a *negative transition (-)* from its own previous move (i.e., rock -> scissors, scissors -> paper, paper -> rock). In other words, the bot usually transitions to the move that will *lose* to what it just played.

* __Opponent previous move (+)__: The bot has a 90% chance of making a *positive transition* from its *opponent's* previous move. In other words, the bot usually transitions to the move that will beat what its opponent just played.

* __Opponent previous move (0)__: The bot has a 90% chance of making a *null transition (0)* from its *opponent's* previous move. In other words, the bot usually copies what its opponent just played.

* __Win-stay-lose-positive__: The bot has a 90% chance of making a null transition after each win, a positive transition after each loss, and a negative transition after each tie. 

* __Win-positive-lose-negative__: The bot has a 90% chance of making a positive transition after each win, a negative transition after each loss, and a null transition after each tie.

* __Outcome-transition dual dependency__: The bot has a 90% chance of making a particular transition (+, -, 0) for every combination of previous outcome *and* previous transition. For example, if the bot has just won from a positive transition, it might have a 90% chance of making a negative transition, but if it just won from a negative transition, it might instead have a 90% chance of a positive transition.

The names for each strategy above are used throughout the remainder of the analyses.


## Results

### Distribution of win count differentials

How well do participants fare against each bot strategy? Shown below is the distribution of participant win count differentials against each bot strategy, with summary statistics labeled for the distribution. 

The expected performance if participants were playing at chance levels is shown in the final plot for comparison: it approximates a Gaussian centered at 0.


```{r bot_distribution_analysis, echo=FALSE}
# Get win count differentials for each bot strategy
wcd_all = get_empirical_win_count_differential(data)
wcd_null = get_null_win_count_differential(10000)

# Generate histograms of each win count differential
hist_prev_move_positive = plot_win_count_differential(wcd_all[wcd_all$bot_strategy == "prev_move_positive",], STRATEGY_LOOKUP[["prev_move_positive"]], -50, NUM_ROUNDS, 50, 5)
hist_prev_move_negative = plot_win_count_differential(wcd_all[wcd_all$bot_strategy == "prev_move_negative",], STRATEGY_LOOKUP[["prev_move_negative"]], -50, NUM_ROUNDS, 50, 5)
hist_opponent_prev_move_positive = plot_win_count_differential(wcd_all[wcd_all$bot_strategy == "opponent_prev_move_positive",], STRATEGY_LOOKUP[["opponent_prev_move_positive"]], -50, NUM_ROUNDS, 200, 7)
hist_opponent_prev_move_nil = plot_win_count_differential(wcd_all[wcd_all$bot_strategy == "opponent_prev_move_nil",], STRATEGY_LOOKUP[["opponent_prev_move_nil"]], -50, NUM_ROUNDS, 50, 4)
hist_win_nil_lose_positive = plot_win_count_differential(wcd_all[wcd_all$bot_strategy == "win_nil_lose_positive",], STRATEGY_LOOKUP[["win_nil_lose_positive"]], -50, NUM_ROUNDS, 150, 5)
hist_win_positive_lose_negative = plot_win_count_differential(wcd_all[wcd_all$bot_strategy == "win_positive_lose_negative",], STRATEGY_LOOKUP[["win_positive_lose_negative"]], -50, NUM_ROUNDS, 180, 4)
hist_outcome_transition_dual_dep = plot_win_count_differential(wcd_all[wcd_all$bot_strategy == "outcome_transition_dual_dependency",], STRATEGY_LOOKUP[["outcome_transition_dual_dependency"]], -50, NUM_ROUNDS, 100, 8)
hist_null = plot_null_win_count_differential(wcd_null, "Chance behavior", -50, 300)
```



```{r bot_distributions, echo=FALSE, fig.width=12, fig.height=12, fig.align="center"}
# Generate plots with patchwork
hist_prev_move_positive + hist_prev_move_negative +
  hist_opponent_prev_move_positive + hist_opponent_prev_move_nil +
  hist_win_nil_lose_positive + hist_win_positive_lose_negative +
  hist_outcome_transition_dual_dep + hist_null +
  plot_layout(ncol = 2, heights = unit(c(4), c("cm")), widths = unit(c(12), c("cm")))
```



### Summary of win count differentials

The distributions above are summarized in the figure below, which shows mean and SEM of empirical win count differentials against each bot strategy. The red line indicates chance performance.


```{r bot_summary_analysis, echo=FALSE}
wcd_summary = get_win_count_differential_summary(wcd_all)
```



```{r bot_summary, echo=FALSE, fig.width=12, fig.height=8, fig.align="center"}
plot_win_count_differential_summary(wcd_summary)
```



### Learning rate against each bot strategy

The above results suggest that participants were able to capitalize early on against the strategies where we see higher win count differentials, but may have struggled to figure out outcome-dependent strategies such as win-stay-lose-positive.


```{r bot_win_pct_analysis, echo=FALSE}
subject_block_data = get_subject_block_data(data, blocksize = BLOCKSIZE)
block_data_summary = get_block_data_summary(subject_block_data)
```


Below, we plot the average of participant win percentages in bins of N rounds over the course of the match (shown below with `r BLOCKSIZE` rounds per bin). The red line indicates chance performance.


```{r bot_win_pct, echo=FALSE, fig.width=12, fig.height=8, fig.alignt="center"}
plot_win_pct_by_block(block_data_summary)
```


Consistent with the overall win count differentials, we see that for the bot strategies that participants achieved a high win count differential against, participants perform well above chance from early on. Meanwhile, for the outcome-dependent strategies (win-stay-lose-positive and win-positive-lose-negative), participants only seem to figure out what the bot is doing in the final third or so of the match. Finally, for the outcome-transition dual dependency strategy, participants never perform above chance levels.



### What do people learn about each bot?

The win percentage data by bin above suggests that people have mostly figured out the dependency structure for the simpler bots from the beginning, but may not fully figure out the outcome-dependent bot strategies: their progress at the end may reflect figuring out single pieces of the bot's dependency structure. 

To investigate this, below we plot overall participant win percentages against each bot strategy, based on the relevant conditioning outcome for that bot strategy. In the first two rows, we show win percentages against the previous move strategies based on bot previous moves and player previous moves, respectively. In the next two rows, we show win percentages against the previous outcome strategies (win, tie, loss) based on the bot's previous outcome, *as well as the outcome two rounds back* for the bot. Though the bot's strategy is only dependent on its previous outcome, people's win count percentages suggest that they fall into patterns of strong dependency on the outcome two rounds ago.


```{r bot_dependency_win_pct_analysis, echo=FALSE}
# 1. Bot previous move strategies
bot_loss_prev_move = get_bot_prev_move_loss_pct(data)
bot_loss_summary_prev_move = get_prev_move_win_pct_summary(bot_loss_prev_move)
  
# Generate plots
prev_move_positive_plot = plot_prev_move_win_pct(bot_loss_summary_prev_move, "prev_move_positive", "Bot previous move")
prev_move_negative_plot = plot_prev_move_win_pct(bot_loss_summary_prev_move, "prev_move_negative", "Bot previous move")

# 2. Player previous move strategies
player_win_prev_move = get_player_prev_move_win_pct(data)
player_win_summary_prev_move = get_prev_move_win_pct_summary(player_win_prev_move)

# Generate plots
opponent_prev_move_positive_plot = plot_prev_move_win_pct(player_win_summary_prev_move, "opponent_prev_move_positive", "Player previous move")
opponent_prev_move_nil_plot = plot_prev_move_win_pct(player_win_summary_prev_move, "opponent_prev_move_nil", "Player previous move")

# 3. Bot previous outcome
bot_loss_prev_outcome = get_bot_prev_outcome_loss_pct(data)
bot_loss_summary_prev_outcome = get_prev_outcome_win_pct_summary(bot_loss_prev_outcome)

# Generate plots
win_nil_lose_positive_plot_outcome = plot_outcome_win_pct(bot_loss_summary_prev_outcome, "win_nil_lose_positive", "Bot previous outcome")
win_positive_lose_negative_plot_outcome = plot_outcome_win_pct(bot_loss_summary_prev_outcome, "win_positive_lose_negative", "Bot previous outcome")


# 4. Bot two outcomes back
bot_loss_prev2_outcome = get_bot_prev2_outcome_loss_pct(data)
bot_loss_summary_prev2_outcome = get_prev_outcome_win_pct_summary(bot_loss_prev2_outcome)

# Generate plots
win_nil_lose_positive_plot_outcome_prev = plot_outcome_win_pct(bot_loss_summary_prev2_outcome, "win_nil_lose_positive", "Bot ante-previous outcome")
win_positive_lose_negative_plot_outcome_prev = plot_outcome_win_pct(bot_loss_summary_prev2_outcome, "win_positive_lose_negative", "Bot ante-previous outcome")


```



```{r bot_dependency_win_pct, echo=FALSE, fig.width=12, fig.height=12, fig.align="center"}
# Plot using patchwork
prev_move_positive_plot + prev_move_negative_plot +
  opponent_prev_move_positive_plot + opponent_prev_move_nil_plot +
  win_nil_lose_positive_plot_outcome + win_positive_lose_negative_plot_outcome +
  win_nil_lose_positive_plot_outcome_prev + win_positive_lose_negative_plot_outcome_prev +
  plot_layout(ncol = 2, heights = unit(c(5), c("cm")), widths = unit(c(12), c("cm")))

```


It's clear that while people have mostly figured out the full dependency for the simpler bots, they may only figure out single aspects of the outcome-dependent bot strategies. Most interestingly of all, people show a modest ability to win against the bot when the previous round was a tie, but a very strong win percentage when the outcome *two rounds ago* was a loss for the bot. 



## Appendix: Slider Data

Below we show Likert scale slider responses for the five questions participants were shown after completing the RPS game. There are no major differences across bot strategies for any of the questions provided. 

```{r slider_analysis, echo=FALSE, message=FALSE}
slider_data = read_slider_data(SLIDER_FILE, data)
slider_summary = get_slider_summary(slider_data)
```



```{r slider_plots, echo=FALSE, fig.width=16, fig.height=20, fig.align="center"}
# Generate plots
slider_qs = unique(slider_summary$statement)

q1_plot = slider_summary %>%
  filter(statement == slider_qs[1]) %>%
  plot_slider_data()

q2_plot = slider_summary %>%
  filter(statement == slider_qs[2]) %>%
  plot_slider_data()

q3_plot = slider_summary %>%
  filter(statement == slider_qs[3]) %>%
  plot_slider_data()

q4_plot = slider_summary %>%
  filter(statement == slider_qs[4]) %>%
  plot_slider_data()

q5_plot = slider_summary %>%
  filter(statement == slider_qs[5]) %>%
  plot_slider_data()


q1_plot + q2_plot + q3_plot + q4_plot + q5_plot +
  plot_layout(ncol = 2)

```

## Appendix: Free Response Data

Below is a table of free response answers, by bot strategy. 
Participants were prompted with the following: *"In the text box below, please describe any strategies you used to try and beat your opponent."*



```{r free_resp, echo=FALSE, message=FALSE}
fr_data = read_free_resp_data(FREE_RESP_FILE, data)

fr_data %>%
  arrange(bot_strategy, strategy, game_id, player_id, free_resp_answer) %>%
  select(strategy, game_id, player_id, free_resp_answer) %>%
  kable(col.names = c("Strategy", "Game ID", "Player ID", "Free Response"))
```







